---
title: R package 'dbpedia' - wrapper for DBpedia Spotlight
output: github_document
editor_options: 
  chunk_output_type: console
---

[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)


## Installation

At this stage, the dbpedia R package is a GitHub-only package. Install it as follows:

```{r install_dbpedia_from_github, eval = FALSE}
devtools::install_github("PolMine/dbpedia", ref = "main")
```

The SPARQL package is a suggested package used internally by functions `` and was archived at CRAN. The latest version (1.16) can be installed as follows.

```{r install_sparql_pkg_from_archive, eval = FALSE}
devtools::install_version("SPARQL", version = "1.16", repos = "http://cran.us.r-project.org") 
```


## Run dockerized DBpedia Spotlight

DBpedia Spotlight docker containers are not available for all architectures. Particularly not for M1. In this case, build container from the dockerfile as follows.

```{sh, eval = FALSE}
git clone https://github.com/dbpedia-spotlight/spotlight-docker.git
cd spotlight-docker
docker build -t dbpedia/dbpedia-spotlight:latest .
```

Then run the image as follows.

```{sh, eval = FALSE}
docker run -tid --restart unless-stopped --name dbpedia-spotlight.de --mount source=spotlight-model,target=/opt/spotlight -p 2222:80  dbpedia/dbpedia-spotlight spotlight.sh de
```


## Using the package

```{r, eval = FALSE}
library(polmineR)
library(dbpedia)

speech <- corpus("GERMAPARL2") %>% 
  subset(protocol_date == "2018-02-22") %>%
  subset(speaker_name == "Angela Merkel") %>%
  subset(p_type == "speech") %>%
  as.speeches(
    s_attribute_name = "speaker_name",
    s_attribute_date = "protocol_date",
    progress = FALSE
  ) %>%
  .[[1L]]

named_entites <- speech %>%
  subset(ne_type) %>%
  as_annotation()

read(speech, annotation = named_entites)

dbpedia_links <- get_dbpedia_links(
  x = speech,
  language = "de",
  mw = "ne_type"
)

read(speech, annotation = dbpedia_links)
```


## Related work

- https://github.com/glaserL/unsc-ne

- https://pypi.org/project/spacy-dbpedia-spotlight/

- https://github.com/news-r/spotlight