---
title: R package 'dbpedia' - wrapper for DBpedia Spotlight
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- badges: start -->
[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
[![R-CMD-check](https://github.com/PolMine/dbpedia/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/PolMine/dbpedia/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

## About

Functionality for entity linking from R: Get DBpedia URIs for existing named entity recognition in a corpus using DBpedia Spotlight.


## Installation

At this stage, the dbpedia R package is a GitHub-only package. Install it as follows:

```{r install_dbpedia_from_github, eval = FALSE}
devtools::install_github("PolMine/dbpedia", ref = "main")
```

The SPARQL package is a suggested package used internally by functions `` and was archived at CRAN. The latest version (1.16) can be installed as follows.

```{r install_sparql_pkg_from_archive, eval = FALSE}
devtools::install_version("SPARQL", version = "1.16", repos = "http://cran.us.r-project.org") 
```


## Run dockerized DBpedia Spotlight

DBpedia Spotlight docker containers are not available for all architectures. Particularly not for M1. In this case, build container from the dockerfile as follows.

```{sh, eval = FALSE}
git clone https://github.com/dbpedia-spotlight/spotlight-docker.git
cd spotlight-docker
docker build -t dbpedia/dbpedia-spotlight:latest .
```

Then run the image as follows.

```{sh, eval = FALSE}
docker run -tid --restart unless-stopped --name dbpedia-spotlight.de --mount source=spotlight-model,target=/opt/spotlight -p 2222:80  dbpedia/dbpedia-spotlight spotlight.sh de
```


## Using the package

Upon loading the dbpedia package, a startup message will issue information whether a Docker container running locally has been detected and on the API and language used.

```{r load_dbpedia, eval = TRUE}
library(dbpedia)
```

This information is available during the R session and is used by the `get_dbpedia_uris()` method.

```{r}
getOption("dbpedia.endpoint")
getOption("dbpedia.lang")
```

```{r, eval = FALSE}
library(polmineR)
library(dplyr)

speech <- corpus("GERMAPARL2") %>% 
  subset(protocol_date == "2018-02-22") %>%
  subset(speaker_name == "Angela Merkel") %>%
  subset(p_type == "speech") %>%
  as.speeches(
    s_attribute_name = "speaker_name",
    s_attribute_date = "protocol_date",
    progress = FALSE
  ) %>%
  .[[1L]]

named_entites <- speech %>%
  subset(ne_type) %>%
  as_annotation()

read(speech, annotation = named_entites)

dbpedia_uris <- get_dbpedia_uris(
  x = speech,
  language = "de",
  s_attribute = "ne_type"
)

read(speech, annotation = as_subcorpus(dbpedia_uris))
```

```{r, eval = FALSE}
library(RCurl)
library(XML)

dbpedia_uris <- dbpedia_uris %>%
  pull(dbpedia_uri) %>% 
  dbpedia_get_wikidata_uris(
    optional = "municipalityCode",
    endpoint = "http://de.dbpedia.org/sparql",
    wait = 0.5,
    limit = 100,
    progress = TRUE
  ) %>% 
  right_join(dbpedia_uris, by = "dbpedia_uri")
```

```{r eval = FALSE}
dbpedia_results <- dbpedia_uris %>%
  pull(wikidata_id) %>% 
  wikidata_query(
    id = "P439", # German municipality key
    wait = 1,
    limit = 100,
    progress = TRUE
  )
```


## Related work

- https://github.com/glaserL/unsc-ne

- https://pypi.org/project/spacy-dbpedia-spotlight/

- https://github.com/news-r/spotlight


## Acknowledgements

We gratefully acknowledge funding from the German National Research Data Infrastructure (Nationale Forschungsdaten-Infrastruktur / NFDI). Developing the dbpedia package is part of the measure "Linking Textual Data" as part of the consortium KonsortSWD (project number 442494171).